{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-12-12 README.Rmd
broke out trend tables into different annual seasonal groups

@@

`trnds_ave.RData` trend summary of averages by annual, monthly categories

`trnds_chg.RData` trend summary of percent changes by annual, monthly categories

`trnds_sk.RData` trend summary using seasonal kendall on flow-normalized results

>>>>

`trnds_seasyr.RData` collection of three data frames for seasonal kendall on observed, fit, norm data, different annual and seasonal periods

@@

### 2016-12-12 sftrends_manu.Rnw
broke out trend tables into different annual seasonal groups

<<trndmap, fig.cap = 'Percent change per year in nitrogen analytes for aggregations by (a) year, (b) seasons from 1976-1995, and (c) seasons from 1996-2014.  Changes are based on seasonal Kendall tests of flow-normalized results within each time period. Station names are shown in the top left panel. Values less than -9\\% change were truncated for visual clarity of the bar plots, see \\cref{tab:trnds} for actual values. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.', fig.height = 8, fig.width = 8, cache = T, out.width = '\\textwidth'>>=

@@

@@

# unnest data frame, assign colors by rank and sign
toplo <- select(trnds_seasyr, Site_Code, resvar, ann, cat, norm) %>% 
  rename(perchg = norm) %>% 
  left_join(., statmeta, by = 'Site_Code') %>% 
  data.frame %>% 
  mutate(
    perchg = pmax(yrng[1], perchg),
    perchg = pmin(yrng[2], perchg),
    negs = sign(perchg),  
    rnks = scales::rescale(abs(perchg), c(1, 1000)), 
    rnks = round(rnks), 
    cols = ifelse(negs == 1, palpos[rnks], palneg[rnks]),
    ind = 1:nrow(.)
    ) %>%
  unite('cattype', cat, ann, sep = '', remove = F) %>%
  mutate(
    cattype = factor(cattype),
    cattype = fct_collapse(cattype, 
      an = c('1976-1995bef', '1996-2014aft'),
      mobef = paste0(c('Spring', 'Summer', 'Fall', 'Winter'), 'bef'), 
      moaft = paste0(c('Spring', 'Summer', 'Fall', 'Winter'), 'aft') 
    )
  )

@@

  # keep bar widths the same for cattypes
  subwid <- 0.1
  if(grepl('an$', uni))
    subwid <- subwid/2

@@

grid.arrange(
  leg,
  arrangeGrob(
    arrangeGrob (plo_ls[[1]], plo_ls[[4]], plo_ls[[7]], ncol = 1, top = c('(a) 1976-1995, 1996-2014')), 
    arrangeGrob(plo_ls[[3]], plo_ls[[6]], plo_ls[[9]], ncol = 1, top = c('(b) Sp, Su, Fa, Wi (1976-1995)')),
    arrangeGrob(plo_ls[[2]], plo_ls[[5]], plo_ls[[8]], ncol = 1, top = c('(c) Sp, Su, Fa, Wi (1996-2014)')),
    ncol = 3
    ),
  ncol = 1, 
  heights = c(0.1, 1)
  )

@@

% overall trends, years
<<trnds>>=
data(trnds_sk)

# percent changes
tab <- trnds_sk %>% 
  select(Site_Code, resvar, cat, med, perchg, ztest) %>% 
  mutate(
    med = round(med, 1), 
    perchg = round(perchg, 1),
    perchg = ifelse(
      perchg < 0, 
      paste0('(', perchg, ')' ), 
      paste0('(\\textit{\\textbf{', perchg, '}})')
      ), 
    ztest = p_ast(ztest)
  ) %>% 
  unite('perchg', perchg, ztest, sep = '') %>% 
  mutate(
    perchg = paste0('\\footnotesize{', perchg, '}')
  ) %>% 
  unite('val', med, perchg, sep = ' ') %>% 
  spread(cat, val) %>% 
  arrange(resvar, Site_Code) %>% 
  select(Site_Code, resvar, `1976-1995`, `1996-2014`, Spring, Summer, Fall, Winter)

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and different time periods.  Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold-italic). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. See \\cref{fig:trndmap} for a summary of spatial trends. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'
foot.val<-'\\footnotesize *$p<0.05$; **$p<0.005$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = c(9, 9, 9),
  cgroup = c('Annual', 'Seasonal'),
  n.cgroup = c(2, 4),
  rowname = stats,
  label = 'tab:trnds',
  insert.bottom = foot.val
  )

>>>>

<<trndmap, fig.cap = 'Percent change per year in nitrogen analytes for aggregations by (a) year, (b) seasons from 1976-1995, and (c) seasons from 1996-2013.  Changes are based on seasonal Kendall tests of flow-normalized results within each time period. Station names are shown in the top left panel. Values less than -9\\% change were truncated for visual clarity of the bar plots, see \\cref{tab:trnds} for actual values. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.', fig.height = 8, fig.width = 8, cache = T, out.width = '\\textwidth'>>=

@@

# combine all
trnds_seasyr <- rbind(trnds_nrm, trnds_fit, trnds_obs) %>% 
  select(Site_Code, resvar, cat, perchg, ann, trndvar) %>% 
  spread(trndvar, perchg) %>% 
  mutate(Site_Code = factor(Site_Code, levels = c('D7', 'D6', 'D4', 'D28', 'D26', 'D19', 'P8', 'C10', 'C3')))

@@

# unnest data frame, assign colors by rank and sign
toplo <- select(trnds_seasyr, Site_Code, resvar, ann, cat, norm) %>% 
  rename(perchg = norm) %>% 
  left_join(., statmeta, by = 'Site_Code') %>% 
  data.frame %>% 
  mutate(
    perchg = pmax(yrng[1], perchg),
    perchg = pmin(yrng[2], perchg),
    negs = sign(perchg),  
    rnks = scales::rescale(abs(perchg), c(1, 1000)), 
    rnks = round(rnks), 
    cols = ifelse(negs == 1, palpos[rnks], palneg[rnks]),
    ind = 1:nrow(.)
    ) %>%
  unite('cattype', cat, ann, sep = '', remove = F) %>%
  mutate(
    cattype = factor(cattype),
    cattype = fct_collapse(cattype, 
      an = c('1976-1995bef', '1996-2013aft'),
      mobef = paste0(c('Spring', 'Summer', 'Fall', 'Winter'), 'bef'), 
      moaft = paste0(c('Spring', 'Summer', 'Fall', 'Winter'), 'aft') 
    )
  )

@@

  # keep bar widths the same for cattypes
  subwid <- 0.1
  if(grepl('an$', uni))
    subwid <- 0.06

@@

grid.arrange(
  leg,
  arrangeGrob(
    arrangeGrob (plo_ls[[1]], plo_ls[[4]], plo_ls[[7]], ncol = 1, top = c('(a) 1976-1995, 1996-2013')), 
    arrangeGrob(plo_ls[[3]], plo_ls[[6]], plo_ls[[9]], ncol = 1, top = c('(b) Sp, Su, Fa, Wi (1976-1995)')),
    arrangeGrob(plo_ls[[2]], plo_ls[[5]], plo_ls[[8]], ncol = 1, top = c('(c) Sp, Su, Fa, Wi (1996-2013)')),
    ncol = 3
    ),
  ncol = 1, 
  heights = c(0.1, 1)
  )

@@

% overall trends, years
<<trndsann>>=
data(trnds_seasyr)

# percent changes
tab <- trnds_nrm %>% 
  filter(cat %in% c('1976-1995', '1996-2013')) %>% 
  select(Site_Code, resvar, cat, med, perchg, ztest) %>% 
  mutate(
    med = round(med, 1), 
    perchg = round(perchg, 1),
    perchg = ifelse(
      perchg < 0, 
      paste0('(', perchg, ')' ), 
      paste0('(\\textit{\\textbf{', perchg, '}})')
      ), 
    ztest = p_ast(ztest)
  ) %>% 
  unite('perchg', perchg, ztest, sep = '') %>% 
  mutate(
    perchg = paste0('\\footnotesize{', perchg, '}')
  ) %>% 
  unite('val', med, perchg, sep = ' ') %>% 
  spread(cat, val) %>% 
  arrange(resvar, Site_Code) %>% 
  select(Site_Code, resvar, `1976-1995`, `1996-2013`)

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and annual aggregations.  Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold-italic). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. See \\cref{fig:trndmap} for a summary of spatial trends.'
foot.val<-'\\footnotesize *$p<0.05$; **$p<0.005$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = c(9, 9, 9),
  cgroup = c('Annual'),
  n.cgroup = c(2),
  rowname = stats,
  label = 'tab:trndsann',
  insert.bottom = foot.val
  )

% overall trends, seasons, first annual periodyears
<<trndsbef>>=
data(trnds_seasyr)

# percent changes
tab <- trnds_nrm %>% 
  filter(
    cat %in% c('Spring', 'Summer', 'Fall', 'Winter') &
    ann %in% 'bef'
    ) %>% 
  select(Site_Code, resvar, cat, med, perchg, ztest) %>% 
  mutate(
    med = round(med, 1), 
    perchg = round(perchg, 1),
    perchg = ifelse(
      perchg < 0, 
      paste0('(', perchg, ')' ), 
      paste0('(\\textit{\\textbf{', perchg, '}})')
      ), 
    ztest = p_ast(ztest)
  ) %>% 
  unite('perchg', perchg, ztest, sep = '') %>% 
  mutate(
    perchg = paste0('\\footnotesize{', perchg, '}')
  ) %>% 
  unite('val', med, perchg, sep = ' ') %>% 
  spread(cat, val) %>% 
  arrange(resvar, Site_Code)

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and seasonal aggregations from 1976-1995. Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold-italic). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. See \\cref{fig:trndmap} for a summary of spatial trends. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'
foot.val<-'\\footnotesize *$p<0.05$; **$p<0.005$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = c(9, 9, 9),
  cgroup = c('Seasonal, 1976-1995'),
  n.cgroup = c(4),
  rowname = stats,
  label = 'tab:trndsbef',
  insert.bottom = foot.val
  )

% overall trends, seasons, first annual periodyears
<<trndsaft>>=
data(trnds_seasyr)

# percent changes
tab <- trnds_nrm %>% 
  filter(
    cat %in% c('Spring', 'Summer', 'Fall', 'Winter') &
    ann %in% 'aft'
    ) %>% 
  select(Site_Code, resvar, cat, med, perchg, ztest) %>% 
  mutate(
    med = round(med, 1), 
    perchg = round(perchg, 1),
    perchg = ifelse(
      perchg < 0, 
      paste0('(', perchg, ')' ), 
      paste0('(\\textit{\\textbf{', perchg, '}})')
      ), 
    ztest = p_ast(ztest)
  ) %>% 
  unite('perchg', perchg, ztest, sep = '') %>% 
  mutate(
    perchg = paste0('\\footnotesize{', perchg, '}')
  ) %>% 
  unite('val', med, perchg, sep = ' ') %>% 
  spread(cat, val) %>% 
  arrange(resvar, Site_Code)

# table stuff
cap.val <- 'Summaries of flow-normalized trends in nitrogen analytes for all stations and seasonal aggregations from 1996-2013. Summaries are  medians (mg L$^{-1}$) and percent change per year in parentheses (increasing in bold-italic). Changes and significance estimates are based on seasonal Kendall tests of flow-normalized results within each time period. See \\cref{fig:trndmap} for a summary of spatial trends. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.'
foot.val<-'\\footnotesize *$p<0.05$; **$p<0.005$'
vars <- c('DIN', 'NH$_{4}^{+}$', 'NO$_{2}^{-}$/NO$_{3}^{2-}$')
stats <- tab$Site_Code

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = c(9, 9, 9),
  cgroup = c('Seasonal, 1996-2013'),
  n.cgroup = c(4),
  rowname = stats,
  label = 'tab:trndsaft',
  insert.bottom = foot.val
  )
